---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# pollspain

<!-- badges: start -->
[![GitHub release](https://img.shields.io/github/v/release/dadosdelaplace/pollspain)](https://github.com/dadosdelaplace/pollspain/releases)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

The main objective of the R package `{pollspain}` is to provide social scientists, political analysts, journalists, and citizens with easy and straightforward [**access to electoral data from Spain**]{.hl}. This includes both **aggregated election results** extracted from polling stations and **electoral survey data** (including housing effects). The package also offers simple tools for seat allocation, vote simulation, and visualization. The package is designed so that users with basic knowledge of `R` can use it, providing [**tidyverse-style functions**]{.hl}.

## Installation

You can [**install the development version**]{.hl} from [GitHub](https://github.com/) with:

```{r eval = FALSE}
# install.packages("devtools") # only if not already installed
devtools::install_github("dadosdelaplace/pollspain")
library(pollspain) # after installing
```


```{r message = FALSE, warning = FALSE, echo = FALSE}
library(pollspain)
```

An **internet connection** is required for the installation, as well as for downloading data. Data is stored in the accompanying `pollspain-data` repository. You can find the **data dictionary** and more information about the data structure at <https://github.com/dadosdelaplace/pollspain-data>

## Usage

### Election summaries

Using `summary_election_data(type_elec = ..., year = ..., level = ...)` with

* a vector of **election dates** (e.g., `year = 2023` or `date = "2023-07-24"`)
* a vector of **election types** (currently, it only works properly for `type_elec = "congress"`)
* a geographic **level for aggregation** (e.g., `level = "ccaa"` or `level = "prov"`)

returns a [**summary table of election results aggregated at the administrative level**]{.hl}. This includes both general data (blank votes, turnout, etc) and the **ballots received by each party or candidacy**. The available [**aggregation levels**]{.hl} (`level`) are: `"all"` (for a national summary), `"ccaa"` (autonomous communities), `"prov"` (province), `"mun"` (municipality), `"mun_district"` (electoral district), `"sec"` (census section), and `"poll_station"`.

```{r summary-election-data-1, eval = FALSE}
# Summary election data at national level for both elections in 2019
# (general data without candidacies ballots)
summary_data_all <- summary_election_data(type_elec = "congress", year = 2019)
summary_data_all
```

```{r summary-election-data-1-b, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
# Summary election data at national level for both elections in 2019
# (general data without candidacies ballots)
summary_data_all <- summary_election_data(type_elec = "congress", year = 2019)
knitr::kable(summary_data_all)
```

```{r summary-election-data-3, eval = FALSE}
# Summary election data, aggregating candidacies ballots at prov level
summary_data_prov_parties <-
  summary_election_data("congress", year = c(2000, 2008, 2023),
                        level = "prov", by_parties = TRUE)
head(summary_data_prov_parties, 5)
```

```{r summary-election-data-3-b, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
# Summary election data, aggregating candidacies ballots at prov level
# for three elections
summary_data_prov_parties <-
  summary_election_data("congress", year = c(2000, 2008, 2023),
                        level = "prov", by_parties = TRUE)
knitr::kable(head(summary_data_prov_parties, 5))
```

`summary_election_data()` is a **user-friendly combination** of the functions `get_election_data()` (which merges different data sources at the polling station level) and `aggregate_election_data()` (which aggregates the data to the requested level). See [**some uses and detailed input arguments**]{.hl} of the function in [...](https://javieralvarezliebana.es/pollspain/articles/...) and [...](https://javieralvarezliebana.es/pollspain/articles/...) for  **advanced users**.

<details>
<summary><strong>⚠️ About municipalities</strong></summary>

The municipality data (names and codes) were **extracted from the version published by the National Statistics Institute (INE) on February 6, 2025**. The configuration of municipalities from previous years has been adapted to the most recent setup, recoding cases where municipalities have merged or disappeared.

Data extracted from <https://www.ine.es/daco/daco42/codmun/codmun20/20codmun.xlsx>

</details>


<details>
<summary><strong>⚠️ About CERA</strong></summary>

According to the National Statistics Institute (INE) «the electoral roll contains the registration of those who meet the requirements to be voters and are not definitively or temporarily deprived of the right to vote. The electoral roll is composed of:

* The electoral roll of Spanish citizens residing in Spain (CER).
* The **electoral roll of Spanish citizens residing abroad (CERA)**.

The electoral roll of residents in Spain who are nationals of countries with Agreements for municipal elections (CERE Agreements), and the electoral roll of citizens of the European Union residing in Spain for municipal and European Parliament elections (CERE EU)».

Los datos relativos a CERA se han agregado a nivel nacional, comunidad autónoma y provincial. ...

</details>


To do:

* explicar bien argumentos y usos
* by_parties y short version por defecto



### Seat allocation


### Resúmenes de encuestas


### Estimación de encuestas


### Simulación de resultados electorales


### Data viz

* barras ordenadas a más a menos (con colores)
* ggparlament
* encuestas + promedio
* barras con resultados + encuestas encima
* barras con % de voto vs %escaños?
* mapa
* ¿algún lollipop para mostrar housing efects? con flechas y eso.


## Other functions

The `{pollspain}` package also provides [**more advanced users with useful functions**]{.hl} to preprocess and analyze electoral data—even their own data, as long as it is provided in a proper format.

* [**Utils**]{.hl}: functions contained in the `utils.R` script are intended to serve as **helper functions for data preprocessing**. See <...> for more examples about how to use them.

```{r}
type_to_code_election(type_elec = "congress")
extract_code("01-04-003-01-004-B", level = "mun")
extract_code("01-04-003-01-004-B", level = "mun", full_cod = TRUE)
```

* [**Import raw data**]]{.hl}: functions starting with `import_..._data()` (code can be found in the `import_elections_data.R` file) are aimed at importing and preprocessing as raw as possible the `.DAT` election files from the Spanish Ministry of Interior files available in the Github repository \url{https://github.com/dadosdelaplace/pollspain-data}. See <...> for more examples about how to use them.


```{r import-poll-station-data, eval = FALSE}
# import and preprocess elections data at poll stations level for given election
# types and dates, providing variables related to turnout, blank/valid votes, etc
poll_data <- import_poll_station_data(type_elec = "congress", year = 2019)
head(poll_data)
```


## Authors

**Javier Álvarez-Liébana (maintainer)**, **David Pereiro-Pol**, **Mafalda González-González**, **Irene Bosque-Gala** and **Mikaela De Smedt**. 

`{pollspain}` package has been part of several Master's Theses from the Master in Computational Data Science at UC3M (Madrid). The package’s usability and functionality have been tested by the following collaborators: 

## References

This package has been designed based on the **following resources and references**

* Albuja J. (2025). R pacakge `{electoral}`: allocating seats methods and party system scores (v0.1.4). <https://cran.r-project.org/web/packages/electoral/index.html>
* García Guzmán P. (2025). WikiBarrio: Explore Spanish socio-demographic data at the neighborhood level. <https://www.wikibarrio.es/>
* García Guzmán P. (2025). ineAtlas: Access to Spanish Household Income Distribution Atlas Data. R package version 0.1.3.9000, <https://github.com/pablogguz/ineAtlas>
* Meleiro H. (2024). infoelectoral: Download Spanish Election Results. R package version 1.0.2, <https://github.com/rOpenSpain/infoelectoral>
* Silge J., Nash J.C., and Graves S. (2018). Navigating the R Package Universe. The R Journal 10 (2): 558–63. <https://doi.org/10.32614/RJ-2018-058>
* Wickham H. and Bryan J. R Packages: Organize, Test, Document, and Share Your Code (2023). <https://r-pkgs.org/>


* Electoral data download from repository of the Spanish Ministry of the Interior. <https://infoelectoral.interior.gob.es/es/elecciones-celebradas/area-de-descargas/>

