---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pollspain

Under development. **Authors**: Javier Álvarez-Liébana, David Pereiro-Pol, Mafalda González-González, Irene Bosque-Gala y Mikaela De Smedt.

## Overview

Pending to detail

## Installation

You can install the development version of R package `{pollspain}` from [GitHub](https://github.com/) through the following code:

```{r eval = FALSE}
install.packages("devtools") # only if not already installed
devtools::install_github("dadosdelaplace/pollspain")
```


```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(pollspain)
```

## Usage

### election data functions

El código de esta familia de funciones se puede encontrar en el script `get_elections_data.R`. La función principal es `summary_election_data()`: dado un vector de fechas electorales, un vector de tipos de elecciones y un nivel geográfico (`"all"`, `"ccaa"`, `"prov"`, ...), la función devuelve al usuario una tabla de resumen de los resultados electorales. Indicando `by_parties = TRUE` la función devuelve los resultados para cada partido o candidatura electoral, al nivel de agregación pedido.

Los niveles de agregación (`level`) pueden ser: `"all"`, `"ccaa"`, `"prov"`, `"mun"`, `"mun_district"` (distrito electoral), `"sec"` (sección censal) y `"poll_station"` (mesa electoral).

```{r summary-election-data-1}
# Summary election data at national level (general data without candidacies ballots)
summary_data_all <- summary_election_data("congress", 2019, 4)
summary_data_all
```

```{r summary-election-data-2}
# Summary election data at ccaa level, aggregating the candidacies ballots
summary_data_ccaa_parties <-
  summary_election_data("congress", 2019, 4, level = "ccaa",
                        by_parties = TRUE)
summary_data_ccaa_parties
```


```{r summary-election-data-3}
# Summary election data at prov level, aggregating the candidacies ballots
summary_data_prov_parties <-
  summary_election_data("congress", 2019, 4, level = "prov",
                        by_parties = TRUE)
knitr::kable(summary_data_prov_parties)
```

Dicha función combina, de manera sencilla para el usuario, `get_election_data()` y `aggregate_election_data()` 



#### get_election_data()


#### aggregate_election_data()

La funciòn `aggregate_election_data()` está destinada a agregar al nivel geográfico pedido la información electoral que ya se le proporcione. 



### import functions

Functions starting with `import_..._data()` (whose code can be found in the `import_elections_data.R` file) aimed at importing and preprocessing as raw as possible the `.DAT` election files from the Spanish Ministry of Interior. All functions download the `.rda` files available in the Github repository \url{https://github.com/dadosdelaplace/pollspain-data} (directly downloaded from MIR website).

#### Census data (level: mun)

`import_mun_census_data()` import municipal census data for one or more elections at the municipal level. This function downloads and processes «raw» municipal data files for specified elections, providing variables related to population, census and administrative codes.

```{r get-mun-census-data}
mun_census_data <- import_mun_census_data("congress", 2019, 4)
mun_census_data
```


#### Poll station data (level: poll station)

`import_poll_station_data()` import and preprocess elections data at poll stations level for given election types and dates, providing variables related to turnout, blank/valid votes, etc


```{r import-poll-station-data}
poll_data <- import_poll_station_data("congress", 2019, c(4, 11))
poll_data
```

#### Candidacies data (level: poll station)

`import_candidacies_data()` import and preprocess candidacies data, providing variables related to id of candidacies, abbreviature and names of them and ballots obtained for each one at each poll station.

```{r import-candidacies-data}
candidacies_data <- import_candidacies_data("congress", 2019, c(4, 11))
candidacies_data
```

#### Datos del CERA (pending)

According to the National Statistics Institute (INE):

«The electoral roll contains the registration of those who meet the requirements to be voters and are not definitively or temporarily deprived of the right to vote. The electoral roll is composed of:

* The electoral roll of Spanish citizens residing in Spain (CER).

* The electoral roll of Spanish citizens residing abroad (CERA).

The electoral roll of residents in Spain who are nationals of countries with Agreements for municipal elections (CERE Agreements), and the electoral roll of citizens of the European Union residing in Spain for municipal and European Parliament elections (CERE EU).»

The `get_CERA_data()` function returns the data related to the CERA.

```{r eval = FALSE}
ccaa_CERA_data <- get_CERA_data(election_data, level = "ccaa")
```




### util functions

The functions contained in the `utils.R` script are intended to serve as helper functions for data preprocessing.

* `type_to_code_election()`: converts from election type (referendum, congress, senate, etc.) to the official ministry code.

```{r}
type_to_code_election(type_elec = "congress")
```

* `extract_code()`: given a polling station code, it returns the code corresponding to the requested level of disaggregation.

```{r}
extract_code("01-04-003-01-004-B", level = "mun")
extract_code("01-04-003-01-004-B", level = "mun", full_cod = TRUE)
```


* `recod_parties()`: recodes party names and their acronyms (pending)
