---
title: "How to use utils functions"
output: rmarkdown::html_vignette
author: "Javier Álvarez Liébana"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{How to use utils functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 7
)
```


This tutorial aims to illustrate how more [**advanced developers can use the helper functions**]{.hl} in the `utils.R` script.

## Required packages

Let's load the required packages for this tutorial.

```{r packages}
library(pollspain)
```

### a

dasdas dsad asdasdasdas dasmdi amsi dasmi masim dasdimqoeqm oqqw qom qm qw mdqomqo mqom dqsom odmqom oqmd oqwmoqm odqwx dasdas dsad asdasdasdas dasmdi amsi dasmi masim dasdimqoeqm oqqw qom qm qw mdqomqo mqom dqsom odmqom oqmd oqwmoqm odqwx dasdas dsad asdasdasdas dasmdi amsi 

```{r}
x <- 1
```


## b


dasdas dsad asdasdasdas dasmdi amsi dasmi masim dasdimqoeqm oqqw qom qm qw mdqomqo mqom dqsom odmqom oqmd oqwmoqm odqwx dasdas dsad asdasdasdas dasmdi amsi dasmi masim dasdimqoeqm oqqw qom qm qw mdqomqo mqom dqsom odmqom oqmd oqwmoqm odqwx dasdas dsad asdasdasdas dasmdi amsi dasmi masim dasdimqoeqm oqqw qom qm qw mdqomqo mqom dqsom odmqom oqmd oqwmoqm odqwx dasdas dsad asdasdasdas dasmdi amsi dasmi masim dasdimqoeqm oqqw qom qm qw mdqomqo mqom dqsom odmqom oqmd oqwmoqm odqwx


```{r}
x <- 1
```


