---
title: "How to use utils functions?"
output: rmarkdown::html_vignette
author: "Javier Álvarez Liébana"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{How to use utils functions?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 7
)
```


This tutorial aims to illustrate how more [**advanced developers can use the helper functions**]{.hl} in the `utils.R` script:

* `type_to_code_election()`
* `extract_code()`
* `recod_mun()`


## Required packages

Let's load the required packages for this tutorial.

```{r packages}
library(pollspain)
```

## Convert type of election to code

According to the data extracted from the [website of the Spanish Ministry of the Interior](https://infoelectoral.interior.gob.es/opencms/es/elecciones-celebradas/area-de-descargas/Spanish), each **type of election is associated in the databases with a specific alphanumeric code**: `"01"` (referendum), `"02"` (congress), `"03"` (senate), `"04"` (local elections), `"06"` (cabildo - Canarian council - elections), and `"07"` (European Parliament elections). However, for users, it's much simpler to provide **lexically explicit terms** — for example, writing `"congress"` for congressional elections or `"senate"` for senatorial elections. The `type_to_code_election()` function allows us to convert the user-provided term into the corresponding code required by our databases.


```{r}
type_to_code_election(type_elec = "congress")
type_to_code_election(type_elec = "senate")
type_to_code_election(type_elec = "local")
```



## Extract region codes

Similarly, many of the package's functions allow the user to **aggregate data at different geographic levels**, which can be easily specified as `"all"`, `"ccaa"`, `"prov"`, `"mun"`, `"mun_district"`, `"sec"` (census tract), and `"poll_station"`. Internally, however, each polling station is geolocated using a set of codes provided by Spain’s National Statistics Institute (INE). The `extract_code()` function enables us to **extract the code corresponding to the requested level of disaggregation from a full polling station code**.

For example, we can apply the function to polling station B of census tract `"004"` within the first municipal district of the municipality of Adra (`"003"`), located in the province of Almería (`"04"`), which belongs to the autonomous community of Andalusia (`"01"`).

```{r}
id_INE_poll_station <- "01-04-003-01-004-B"
```

What would be the municipality code corresponding to the polling station?


```{r}
extract_code(id_INE_poll_station, level = "mun")
```

What about the province code?

```{r}
extract_code(id_INE_poll_station, level = "prov")
```

What about the municipal district?


```{r}
extract_code(id_INE_poll_station, level = "mun_district")
```

By default, it only returns the code at the requested aggregation level, but if you set `full_cod = TRUE`, you can obtain the full code, including not only the code for that level but also for all higher levels. For example, every province has a municipality with the code `"001"`, so we need the corresponding province and autonomous community codes to uniquely identify it.

```{r}
extract_code(id_INE_poll_station, level = "mun", full_cod = TRUE)
extract_code(id_INE_poll_station, level = "prov", full_cod = TRUE)
extract_code(id_INE_poll_station, level = "mun_district", full_cod = TRUE)
```

## Recoding municipality names

Over the years, various municipal mergers have taken place in Spain, which means that not all elections feature the same set of municipalities or the same identifying codes. In order to unify and standardize the results provided to users, all output tables refer to the most recent municipality recoding by the Spanish National Statistics Institute (INE) (latest update: 2023). The helper function `recod_mun()` transparently returns the updated code based on that recoding, reassigning codes for merged municipalities accordingly.
